---
title: "EDA: Numeric Summaries"
author: "Andy Turner"
date: "2023-10-05"
output: github_document
---

**group_by()** makes grouping explicit and adds a layer to yoru data

* Based on existing variables 
* Changes behavior of some key functions 
* Not exactly invisible, but it’s easy to miss … 

**summarize()** allows you to compute one-number summaries

* Based on existing variables
* Most useful in conjunction with group_by() 
* Produces a dataframe with grouping variables and summaries –Easy to integrate into a pipeline 



*Basic Graph Components**

* Data: our actual information 

* Aesthetic mappings: Need to map the aesthetics of the plot to the variables --> need to know what we actually want

* geoms: A layer combines data, aesthetic mapping, a geom (geometric object)

## Viz and EDA
```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(lubridate)

knitr::opts_chunk$set(
  fig.width= 6,
  fig.asp= .6,
  out.width= "90%"
)
```

Pulling in data from a national weather registry. Keeping certain variables and recoding the name to be more descriptive in our case. 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())

```

## Initial numeric work

```{r}
weather_df |> 
  ggplot(aes(x=prcp))+
  geom_histogram()
```

here are the big outliers
```{r}
weather_df |> 
  filter(prcp >1000)
```

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x=tmin, y=tmax, color= name))+
  geom_point()
```

## Grouping

```{r}
weather_df |> 
  group_by(name)
```
^ Grouping often looks invisible because there isn't anything that shows up in environment; however, there is a line that shows up in the console that says: "Groups:   name [3]"

```{r}
weather_df |> 
  group_by(name, month)
```
 Groups:   name, month [72] <- 72 unique groups in the data set
 
## Counting

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(n_obs= n())
```
- Around 30 in each group (which makes sense since 1 a day)

```{r}
weather_df |> 
  group_by(month) |> 
  summarize(n_obs= n())
```
- Around 90 in each group (tracks as there are 3 stations each 1 a day)

```{r}
weather_df |> 
  group_by(name) |> 
  summarize(n_obs= n())
```
- 730 each (2 years of data!)

```{r}
weather_df |> 
  count(name, month) |> 
  pivot_wider(
    names_from= name,
    values_from= n
  )
```

## General Summaries

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm= TRUE ),
    median_tmax= median(tmax, na.rm= TRUE),
    sd_tmax= sd(tmax, na.rm=TRUE)
  )
```
`na.rm` is to remove missing values/NAs!!! `na.rm = TRUE` would remove the values. 

When we are using summarize, we can make a temporary data frame using grouping and summarizing

```{r}
weather_df |> 
  group_by(name,month) |> 
  summarize(mean_tmax= mean(tmax, na.rm=TRUE)) |> 
  ggplot(aes(x= month, y=mean_tmax, color=name))+
  geom_point()+
  geom_line()
```
We can use the grouped together or summarized collections and push it right into a graph. 

```{r}
weather_df |> 
  group_by(name,month) |> 
  summarize(mean_tmax= mean(tmax, na.rm=TRUE)) |> 
  pivot_wider(
    names_from= name,
    values_from= mean_tmax
  ) |> 
  knitr::kable(digits=2)
```
Using `knitr: kable()`formats everything as a table we would want for a DF.
